### 2023-03-14 appendix.Rmd
revisions #4

>>>>

```{r echo=F, warning=F, message=F}
loadd(cfa4)
loadd(cfa5)
loadd(sem9)
```

### Confirmatory factor analyses and structural equation models with robust estimator

We refitted our main models using a robust estimator (weighted least square mean and variance adjusted estimator) to ensure that our inferences are not biased due to non-normal and ordinal response distributions.

When fitted with this robust estimator, our first confirmatory factor model fitted the data well (RMSEA = `r printnum(as.numeric(fitMeasures(cfa4)['rmsea']))`; SRMR = `r printnum(as.numeric(fitMeasures(cfa4)['srmr']))`; CFI = `r printnum(as.numeric(fitMeasures(cfa4)['cfi']))`). This model loaded cooperative behaviour in the Dictator Game, the Trust Game, and the Public Goods Game onto the "cooperation" latent variable, and punishment of defectors in the Ultimatum Game, Third-Party Punishment Game, and Second-Party Punishment Game onto the "punishment" latent variable. As in our main model, the two latent variables were significantly negatively correlated with one another (_r_ = `r printnum(standardizedSolution(cfa4)[25,4])`, _p_ = `r printp(parameterEstimates(cfa4)[25,7])`). Model fit improved further when additionally loading the Stag Hunt Game onto the cooperation latent variable and loading the Stag Hunt Game with Punishment (Punish Defector) onto the punishment latent variable (RMSEA = `r printnum(as.numeric(fitMeasures(cfa5)['rmsea']))`; SRMR = `r printnum(as.numeric(fitMeasures(cfa5)['srmr']))`; CFI = `r printnum(as.numeric(fitMeasures(cfa5)['cfi']))`).

We additionally fitted our main structural equation model, including SDO and RWA as predictors of the two latent variables, with the robust estimator. Replicating the result of our main model, we found that SDO negatively predicted the cooperation latent variable (standardised $\beta$ = `r printnum(standardizedSolution(sem9)[19,5])`, unstandardised _b_ = `r printnum(parameterEstimates(sem9)[19,5])`, 95% confidence interval [`r printnum(parameterEstimates(sem9)[19,9])`, `r printnum(parameterEstimates(sem9)[19,10])`], _p_ `r printp(parameterEstimates(sem9)[19,8])`) and positively predicted the punishment latent variable (standardised $\beta$ = `r printnum(standardizedSolution(sem9)[21,5])`, unstandardised _b_ = `r printnum(parameterEstimates(sem9)[21,5])`, 95% confidence interval [`r printnum(parameterEstimates(sem9)[21,9])`, `r printnum(parameterEstimates(sem9)[21,10])`], _p_ = `r printp(parameterEstimates(sem9)[21,8])`). By contrast, RWA was unrelated to both the cooperation latent variable (standardised $\beta$ = `r printnum(standardizedSolution(sem9)[20,5])`, unstandardised _b_ = `r printnum(parameterEstimates(sem9)[20,5])`, 95% confidence interval [`r printnum(parameterEstimates(sem9)[20,9])`, `r printnum(parameterEstimates(sem9)[20,10])`], _p_ = `r printp(parameterEstimates(sem9)[20,8])`) and the punishment latent variable (standardised $\beta$ = `r printnum(standardizedSolution(sem9)[22,5])`, unstandardised _b_ = `r printnum(parameterEstimates(sem9)[22,5], digits = 3)`, 95% confidence interval [`r printnum(parameterEstimates(sem9)[22,9], digits = 3)`, `r printnum(parameterEstimates(sem9)[22,10], digits = 3)`], _p_ = `r printp(parameterEstimates(sem9)[22,8])`).

### 2023-03-14 manuscript.Rmd
revisions #4

abstract: |
 Decades of research suggest that our political differences are best captured by two dimensions of political ideology. The dual evolutionary framework of political ideology predicts that these dimensions should be related to variation in social preferences for cooperation and group conformity. Here, we combine data from a New Zealand survey and a suite of incentivised behavioural tasks (_n_ = 991) to test whether cooperative and conformist preferences covary with a pair of widely used measures of the two dimensions of political ideology --- Social Dominance Orientation (SDO) and Right Wing Authoritarianism (RWA) --- and related policy views. As predicted, we find that cooperative behaviour covaries with lower SDO and economically progressive policy views, while conformist behaviour in the form of social information use covaries with higher RWA and socially conservative policy views. However, we did not find the predicted relationships between punitive and rule following behaviours and RWA or socially conservative views, raising questions about the interpretation of punishment and rule following tasks and the nature of authoritarian conformist preferences. These findings reveal how cooperative and conformist preferences that evolved to help us navigate social challenges in our ancestral past continue to track our political differences even today.

@@

Since political attitudes have a non-trivial heritable component[@Hatemi2014], it is possible that the two dimensions of political ideology have a partly biological basis. Indeed, evolutionary theorists have attempted to explain the existence of the two dimensions in various ways. For example, political egalitarianism has been shown to covary with upper-body strength and physical dominance [@Petersen2019], whereas authoritarianism and traditionalism have been linked to the behavioural immune system [@Tybur2016], negativity bias [@Hibbing2014], and adherence to social norms [@Kessler2008]. Moreover, the two dimensions of ideology have distinct psychological correlates. Studies have shown that economic conservatism covaries with agreeableness[@Heaven2001] and Machiavellianism[@Jones2013], while social conservatism covaries with openness to experience[@Duckitt2010] and need for closure[@Jost2003]. But while these previous theories and studies have illuminated aspects of each dimension separately, they do not attempt to explain the existence of a two-dimensional ideological structure from first principles.

@@

Previous research has studied the covariation between behaviour in incentivised tasks and the two dimensions of political ideology. A number of studies find support for an association between cooperative preferences and SDO, an ideological measure of preferences for hierarchy and anti-egalitarianism [@Haesevoets2015; @Halali2018]. A recent meta-analysis of the links between personality and prosocial behaviour found that cooperative behaviour in a number of social dilemma games that allow exploitation of others was negatively correlated with SDO[@Thielmann2020]. In contrast, other research has failed to find a relationship between SDO and cooperative behaviour in the Prisoner's Dilemma Game, in which two players choose whether to cooperate or defect [@Balliet2016; @Haesevoets2015], or the Stag Hunt Game, in which two players must coordinate on one of multiple equilibrium strategies [@Haesevoets2015]. These findings suggest that SDO may relate to particular affordances of individual games (e.g., exploitation[@Thielmann2020]) rather than to a general behavioural cooperative preference that applies across situations with different payoff structures, as predicted by the dual evolutionary framework.

@@

All games involved one-shot decisions between multiple anonymous players. Such one-shot anonymous games remove the well-documented effects of communication and known players on behaviour [@Chaudhuri2009; @Chaudhuri2011] and avoid potential confounds arising from repeated play, given that repeated game play may result in "super-games" and consequent cooperation even in finite repetitions [@Kreps1982]. The strategy method, in which participants provide responses for all possible roles in the games, was used to measure responses. This approach avoids feedback from other players, allowing us to make claims about cooperative preferences. The strategy method has been shown to elicit similar behaviour to sequential play [@Brandts2000].

@@

Participants were paid a fixed \$20 NZD show-up fee, which ensured at least the living wage payment for a 50 minute study session. This show-up fee was important to make sure that no participant felt disappointed with the study, thereby leading them to drop out of the larger New Zealand Attitudes and Values Study. In addition, participants could earn a bonus payment of between \$10 - 35 (_M_ = \$`r printnum(mean(d1$egame.Paid.T10-20))`, _SD_ = \$`r printnum(sd(d1$egame.Paid.T10-20))`) depending on the decisions made by them and their peers. Although these bonus payments are relatively flat when compared to the show-up fee, previous work has shown that relatively low stakes do not cause marked changes in behaviour[@Cameron1999]. Nevertheless, it should be noted that the absolute value of our bonus payment is higher than that provided in many behavioural economics studies. On average, participants took `r printnum(mean(d1$egame.SecsAll.T10 / 60),digits=0)` minutes to complete the eight economic games (_SD_ = `r printnum(sd(d1$egame.SecsAll.T10 / 60))` mins, range = `r printnum(min(d1$egame.SecsAll.T10 / 60),digits=0)` - `r printnum(max(d1$egame.SecsAll.T10 / 60),digits=0)` mins).

@@

Before examining the relationship between gameplay and ideology, we first tested whether behaviour across the games shows a two-factor structure in line with our pre-registered expectations. Distributions of game decisions are reported as histograms in Supplementary Figure \@ref(fig:gamesDistPlot). Pairwise Spearman's rank correlations between game decisions, accounting for multiple comparisons using Benjamini-Hochberg corrected _p_ values, are visualised in Figure \@ref(fig:corPcaPlot)a. Replicating previous work[@Peysakhovich2014], we found significant positive correlations between most cooperation decisions ($\rho$ = `r printnum(min(cor$r[1:5,1:5][lower.tri(cor$r[1:5,1:5])]))`-`r printnum(max(cor$r[1:5,1:5][lower.tri(cor$r[1:5,1:5])]))`), significant positive correlations between all punishment decisions ($\rho$ = `r printnum(min(cor$r[6:9,6:9][lower.tri(cor$r[6:9,6:9])]))`-`r printnum(max(cor$r[6:9,6:9][lower.tri(cor$r[6:9,6:9])]))`), and non-significant or negative correlations between most cooperation and punishment decisions. Principal components analyses suggested that these patterns of covariation could be explained by two underlying factors: cooperation and punishment. When including only games from previous work, principal components analysis using orthogonal varimax rotation supported a two-factor solution, based on the point of inflexion in the scree plot and the number of factors with eigenvalues above one[@Field2012;@Kaiser1960] (Supplementary Figure \@ref(fig:screePlot)a). This analysis revealed that cooperative behaviour loads highly onto one factor and punitive behaviour loads highly onto a second distinct factor (Figure \@ref(fig:corPcaPlot)b). Together, these factors explained `r printnum(pca1.2$Vaccounted[3,2]*100, digits=0)`% of the variance in game decisions. The two-factor structure held when adding our novel coordination games (Figure \@ref(fig:corPcaPlot)c and Supplementary Figure \@ref(fig:screePlot)b; `r printnum(pca2.2$Vaccounted[3,2]*100, digits = 0)`% variance explained). This factor structure was also supported by exploratory parallel analysis, which independently proposed a `r english(readd(parallel)$nfac)`-factor solution (Supplementary Figure \@ref(fig:parallelPlot)), and confirmatory factor analyses, which were robust to the inclusion of method factors (see Supplementary Results).

@@

Having confirmed the existence of a two-factor structure underlying gameplay, we next fitted pre-registered structural equation models to test the relationships between the two latent variables and political ideology. For reference, previous work has shown that effect sizes for relationships between economic game behaviour and self-reported attitudes tend to vary from very small to medium[@Thielmann2020] (0.00 $\leq$ _r_ $\leq$ 0.26).

We simultaneously regressed the cooperation and punishment latent variables onto mean scores of six Likert-scale items for SDO (Cronbach's $\alpha$ = `r printnum(readd(alphaSDO))`, $\omega$ = `r printnum(omega$FactorLevelIndices$Omega[2])`) and RWA ($\alpha$ = `r printnum(readd(alphaRWA))`, $\omega$ = `r printnum(omega$FactorLevelIndices$Omega[3])`). In line with our second pre-registered hypothesis, we found that SDO significantly negatively predicted the cooperation latent variable with a small-to-medium effect size (standardised $\beta$ = `r printnum(standardizedSolution(sem1)[19,5])`, unstandardised _b_ = `r printnum(parameterEstimates(sem1)[19,5])`, 95% confidence interval [`r printnum(parameterEstimates(sem1)[19,9])`, `r printnum(parameterEstimates(sem1)[19,10])`], _p_ `r printp(parameterEstimates(sem1)[19,8])`, semi-partial _r_ = `r printnum(r_SDO_coop)`; Figure \@ref(fig:coopPunPlot)a). SDO also significantly positively predicted the punishment latent variable with a small effect size, which we did not hypothesise ($\beta$ = `r printnum(standardizedSolution(sem1)[21,5])`, _b_ = `r printnum(parameterEstimates(sem1)[21,5], digits=3)`, 95% CI [`r printnum(parameterEstimates(sem1)[21,9], digits=3)`, `r printnum(parameterEstimates(sem1)[21,10], digits=3)`], _p_ = `r printp(parameterEstimates(sem1)[21,8])`, semi-partial _r_ = `r printnum(r_SDO_pun)`), though the magnitude of this association was not as strong as with the cooperation latent variable (difference in absolute unstandardized coefficients = `r printnum(parameterEstimates(sem1)[59,5], digits=2)`, SE = `r printnum(parameterEstimates(sem1)[59,6], digits=3)`, p `r printp(parameterEstimates(sem1)[59,8])`). In contrast with our third pre-registered hypothesis, we found no significant relationship between RWA and the punishment latent variable ($\beta$ = `r printnum(standardizedSolution(sem1)[22,5])`, _b_ = `r printnum(parameterEstimates(sem1)[22,5], digits=3)`, 95% CI [`r printnum(parameterEstimates(sem1)[22,9], digits=3)`, `r printnum(parameterEstimates(sem1)[22,10], digits=3)`], _p_ = `r printp(parameterEstimates(sem1)[22,8])`, semi-partial _r_ = `r printnum(r_RWA_pun)`; Figure \@ref(fig:coopPunPlot)b). RWA was also unrelated to the cooperation latent variable ($\beta$ = `r printnum(standardizedSolution(sem1)[20,5])`, _b_ = `r printnum(parameterEstimates(sem1)[20,5], digits=2)` [`r printnum(parameterEstimates(sem1)[20,9], digits=2)`, `r printnum(parameterEstimates(sem1)[20,10], digits=2)`], _p_ = `r printp(parameterEstimates(sem1)[20,8])`, semi-partial _r_ = `r printnum(r_RWA_coop)`). This pattern of results was robust to method bias (Supplementary Results) and held when additionally controlling for age, gender, ethnicity, education, socio-economic status, local deprivation, and religiosity (Supplementary Table \@ref(tab:semTable)).

@@

By contrast, the punishment latent variable did not predict social views like support for same-sex marriage (_b_ = `r printnum(parameterEstimates(policy04)[20,4], digits=2)`, 95% CI [`r printnum(parameterEstimates(policy04)[20,8], digits=2)`, `r printnum(parameterEstimates(policy04)[20,9], digits=2)`], _p_ = `r printp(parameterEstimates(policy04)[20,7])`, semi-partial _r_ = `r printnum(r_policy04_pun, digits=2)`), euthanasia (_b_ = `r printnum(parameterEstimates(policy05)[20,4], digits=2)`, 95% CI [`r printnum(parameterEstimates(policy05)[20,8], digits=2)`, `r printnum(parameterEstimates(policy05)[20,9], digits=2)`], _p_ = `r printp(parameterEstimates(policy05)[20,7])`, semi-partial _r_ = `r printnum(r_policy05_pun, digits=2)`), or abortion (_b_ = `r printnum(parameterEstimates(policy06)[20,4], digits=2)`, 95% CI [`r printnum(parameterEstimates(policy06)[20,8], digits=2)`, `r printnum(parameterEstimates(policy06)[20,9], digits=2)`], _p_ `r printp(parameterEstimates(policy06)[20,7])`, semi-partial _r_ = `r printnum(r_policy06_pun, digits=2)`). The punishment latent variable was also unrelated to a latent variable that captured progressive views on all these social policies combined ($\beta$ = `r printnum(standardizedSolution(policyLatent)[29,4])`, _b_ = `r printnum(parameterEstimates(policyLatent)[29,4])`, 95% CI [`r printnum(parameterEstimates(policyLatent)[29,8])`, `r printnum(parameterEstimates(policyLatent)[29,9])`], _p_ = `r printp(parameterEstimates(policyLatent)[29,7])`, semi-partial _r_ = `r printnum(sqrt(inspect(policyLatent, "r2")[["soc"]]))`).

@@

However, it is also possible that our norm-enforcing punishment and rule following tasks were poor measures of a general group conformist preference. As noted above, these tasks contain competitive and selfish elements that may explain their correlations with SDO rather than RWA. The tasks did not explicitly measure social norms [@Krupka2013] or allow players to interact over multiple rounds, limiting the potential for the emergence and enforcement of normative expectations. The tasks were also abstract and context-neutral: they did not specify who the players in the interaction were or contain any differentiation between in-groups and out-groups. This avoids conflating participants' general predisposition to conform with their identification with a particular group, but may fail to activate conformist preferences that depend on specific group contexts, such as socially conservative group identities [@Duckitt2009;@Everett2013]. Such an explanation would imply that authoritarians and social conservatives conform to social information in all contexts, but will only enforce and adhere to norms in specific group contexts. Future research could adapt the methods used here to further refine our understanding of the behavioural preferences underlying the conformist dimension of political ideology.

>>>>

abstract: |
 Decades of research suggest that our political differences are best captured by two dimensions of political ideology. The dual evolutionary framework of political ideology predicts that these dimensions should be related to variation in social preferences for cooperation and group conformity. Here, we combine data from a New Zealand survey and a suite of incentivised behavioural tasks (_n_ = 991) to test whether cooperative and conformist preferences covary with a pair of widely used measures of the two dimensions of political ideology --- Social Dominance Orientation (SDO) and Right Wing Authoritarianism (RWA) --- and related policy views. As predicted, we find that cooperative behaviour is negatively related to SDO and economically conservative policy views, while conformist behaviour in the form of social information use is positively related to RWA and socially conservative policy views. However, we did not find the predicted relationships between punitive and rule following behaviours and RWA or socially conservative views, raising questions about the interpretation of punishment and rule following tasks and the nature of authoritarian conformist preferences. These findings reveal how cooperative and conformist preferences that evolved to help us navigate social challenges in our ancestral past continue to track our political differences even today.

@@

Since political attitudes have a non-trivial heritable component[@Hatemi2014], it is possible that the two dimensions of political ideology have a partly biological basis. Indeed, evolutionary theorists have attempted to explain the existence of the two dimensions in various ways. For example, political egalitarianism has been shown to covary with upper-body strength and physical dominance [@Petersen2019], whereas authoritarianism and traditionalism have been linked to the behavioural immune system [@Tybur2016], negativity bias [@Hibbing2014], and adherence to social norms [@Kessler2008]. Moreover, the two dimensions of ideology have distinct psychological correlates. Studies have shown that economic conservatism covaries with agreeableness[@Heaven2001] and Machiavellianism[@Jones2013], while social conservatism covaries with openness to experience[@Duckitt2010a] and need for closure[@Jost2003]. But while these previous theories and studies have illuminated aspects of each dimension separately, they do not attempt to explain the existence of a two-dimensional ideological structure from first principles.

@@

Previous research has studied the covariation between behaviour in incentivised tasks and the two dimensions of political ideology. A number of studies find support for an association between cooperative preferences and SDO, an ideological measure of preferences for hierarchy and anti-egalitarianism [@Haesevoets2015; @Halali2018]. A recent meta-analysis of the links between personality and prosocial behaviour found that cooperative behaviour in a number of social dilemma games that allow exploitation of others was negatively correlated with SDO[@Thielmann2020]. In contrast, other research has failed to find a relationship between SDO and cooperative behaviour in the Prisoner's Dilemma Game, in which two players choose whether to cooperate or defect [@Balliet2016; @Haesevoets2015], or the Stag Hunt Game, in which two players must coordinate on one of multiple equilibrium strategies [@Haesevoets2015]. These findings suggest that SDO may relate to particular affordances of individual games (e.g., exploitation[@Thielmann2020]) rather than to a general cooperative preference that applies across situations with different payoff structures, as predicted by the dual evolutionary framework.

@@

All games involved one-shot decisions between multiple anonymous players. Such one-shot anonymous games remove the well-documented effects of communication, siganalling, and/or reputation building on behaviour [@Chaudhuri2009; @Chaudhuri2011] and avoid potential confounds arising from repeated play, given that repeated game play may result in "super-games" and consequent cooperation even in finite repetitions [@Kreps1982]. The strategy method, in which participants provide responses for all possible roles in the games, was used to measure responses. This approach avoids feedback from other players, allowing us to make claims about cooperative preferences. The strategy method has been shown to elicit similar behaviour to sequential play [@Brandts2000].

@@

Participants were paid a fixed \$20 NZD show-up fee, which ensured at least the living wage payment for a 50 minute study session. This show-up fee was important to make sure that no participant felt disappointed with the study, thereby leading them to drop out of the larger New Zealand Attitudes and Values Study. In addition, participants could earn a bonus payment of between \$10 - 35 (_M_ = \$`r printnum(mean(d1$egame.Paid.T10-20))`, _SD_ = \$`r printnum(sd(d1$egame.Paid.T10-20))`) depending on the decisions made by them and their peers. While these bonus payments were comparable in size to the show-up fee, previous work has shown that even relatively low stakes do not cause marked changes in behaviour[@Cameron1999]. Nevertheless, it should be noted that the absolute value of our bonus payment is higher than that provided in many behavioural economics studies. On average, participants took `r printnum(mean(d1$egame.SecsAll.T10 / 60),digits=0)` minutes to complete the eight economic games (_SD_ = `r printnum(sd(d1$egame.SecsAll.T10 / 60))` mins, range = `r printnum(min(d1$egame.SecsAll.T10 / 60),digits=0)` - `r printnum(max(d1$egame.SecsAll.T10 / 60),digits=0)` mins).

@@

Before examining the relationship between gameplay and ideology, we first tested whether behaviour across the games shows a two-dimensional structure in line with our pre-registered expectations. Distributions of game decisions are reported as histograms in Supplementary Figure \@ref(fig:gamesDistPlot). Pairwise Spearman's rank correlations between game decisions, accounting for multiple comparisons using Benjamini-Hochberg corrected _p_ values, are visualised in Figure \@ref(fig:corPcaPlot)a. Replicating previous work[@Peysakhovich2014], we found significant positive correlations between most cooperation decisions ($\rho$ = `r printnum(min(cor$r[1:5,1:5][lower.tri(cor$r[1:5,1:5])]))`-`r printnum(max(cor$r[1:5,1:5][lower.tri(cor$r[1:5,1:5])]))`), significant positive correlations between all punishment decisions ($\rho$ = `r printnum(min(cor$r[6:9,6:9][lower.tri(cor$r[6:9,6:9])]))`-`r printnum(max(cor$r[6:9,6:9][lower.tri(cor$r[6:9,6:9])]))`), and non-significant or negative correlations between most cooperation and punishment decisions. Principal components analyses suggested that these patterns of covariation could be explained by two underlying principal components: cooperation and punishment. When including only games from previous work, principal components analysis using orthogonal varimax rotation supported a two-dimensional solution, based on the point of inflexion in the scree plot and the number of principal components with eigenvalues above one[@Field2012;@Kaiser1960] (Supplementary Figure \@ref(fig:screePlot)a). This analysis revealed that cooperative behaviour loads highly onto one principal component and punitive behaviour loads highly onto a second distinct principal component (Figure \@ref(fig:corPcaPlot)b). Together, these factors explained `r printnum(pca1.2$Vaccounted[3,2]*100, digits=0)`% of the variance in game decisions. The two-dimensional structure held when adding our novel coordination games (Figure \@ref(fig:corPcaPlot)c and Supplementary Figure \@ref(fig:screePlot)b; `r printnum(pca2.2$Vaccounted[3,2]*100, digits = 0)`% variance explained). This structure was also supported by exploratory parallel analysis, which independently proposed a `r english(readd(parallel)$nfac)`-dimensional solution (Supplementary Figure \@ref(fig:parallelPlot)), and confirmatory factor analyses, which were robust to the inclusion of method factors (see Supplementary Results).

@@

Having confirmed the existence of a two-dimensional structure underlying gameplay, we next fitted pre-registered structural equation models to test the relationships between the two latent variables and political ideology. For reference, previous work has shown that effect sizes for relationships between economic game behaviour and self-reported attitudes tend to vary from very small to medium[@Thielmann2020] (0.00 $\leq$ _r_ $\leq$ 0.26).

We simultaneously regressed the cooperation and punishment latent variables onto mean scores of six Likert-scale items for SDO (Cronbach's $\alpha$ = `r printnum(readd(alphaSDO))`, $\omega$ = `r printnum(omega$FactorLevelIndices$Omega[2])`) and RWA ($\alpha$ = `r printnum(readd(alphaRWA))`, $\omega$ = `r printnum(omega$FactorLevelIndices$Omega[3])`). In line with our second pre-registered hypothesis, we found that SDO significantly negatively predicted the cooperation latent variable with a small-to-medium effect size (standardised $\beta$ = `r printnum(standardizedSolution(sem1)[19,5])`, unstandardised _b_ = `r printnum(parameterEstimates(sem1)[19,5])`, 95% confidence interval [`r printnum(parameterEstimates(sem1)[19,9])`, `r printnum(parameterEstimates(sem1)[19,10])`], _p_ `r printp(parameterEstimates(sem1)[19,8])`, semi-partial _r_ = `r printnum(r_SDO_coop)`; Figure \@ref(fig:coopPunPlot)a). SDO also significantly positively predicted the punishment latent variable with a small effect size, which we did not hypothesise ($\beta$ = `r printnum(standardizedSolution(sem1)[21,5])`, _b_ = `r printnum(parameterEstimates(sem1)[21,5], digits=3)`, 95% CI [`r printnum(parameterEstimates(sem1)[21,9], digits=3)`, `r printnum(parameterEstimates(sem1)[21,10], digits=3)`], _p_ `r printp(parameterEstimates(sem1)[21,8])`, semi-partial _r_ = `r printnum(r_SDO_pun)`), though the magnitude of this association was not as strong as with the cooperation latent variable (difference in absolute unstandardized coefficients = `r printnum(parameterEstimates(sem1)[59,5], digits=2)`, SE = `r printnum(parameterEstimates(sem1)[59,6], digits=3)`, p `r printp(parameterEstimates(sem1)[59,8])`). In contrast with our third pre-registered hypothesis, we found no significant relationship between RWA and the punishment latent variable ($\beta$ = `r printnum(standardizedSolution(sem1)[22,5])`, _b_ = `r printnum(parameterEstimates(sem1)[22,5], digits=3)`, 95% CI [`r printnum(parameterEstimates(sem1)[22,9], digits=3)`, `r printnum(parameterEstimates(sem1)[22,10], digits=3)`], _p_ = `r printp(parameterEstimates(sem1)[22,8])`, semi-partial _r_ = `r printnum(r_RWA_pun)`; Figure \@ref(fig:coopPunPlot)b). RWA was also unrelated to the cooperation latent variable ($\beta$ = `r printnum(standardizedSolution(sem1)[20,5])`, _b_ = `r printnum(parameterEstimates(sem1)[20,5], digits=2)`, 95% CI [`r printnum(parameterEstimates(sem1)[20,9], digits=2)`, `r printnum(parameterEstimates(sem1)[20,10], digits=2)`], _p_ = `r printp(parameterEstimates(sem1)[20,8])`, semi-partial _r_ = `r printnum(r_RWA_coop)`). This pattern of results was robust to method bias (Supplementary Results) and held when additionally controlling for age, gender, ethnicity, education, socio-economic status, local deprivation, and religiosity (Supplementary Table \@ref(tab:semTable)).

@@

By contrast, the punishment latent variable did not predict social views like support for same-sex marriage (_b_ = `r printnum(parameterEstimates(policy04)[20,4], digits=2)`, 95% CI [`r printnum(parameterEstimates(policy04)[20,8], digits=2)`, `r printnum(parameterEstimates(policy04)[20,9], digits=2)`], _p_ = `r printp(parameterEstimates(policy04)[20,7])`, semi-partial _r_ = `r printnum(r_policy04_pun, digits=2)`), euthanasia (_b_ = `r printnum(parameterEstimates(policy05)[20,4], digits=2)`, 95% CI [`r printnum(parameterEstimates(policy05)[20,8], digits=2)`, `r printnum(parameterEstimates(policy05)[20,9], digits=2)`], _p_ = `r printp(parameterEstimates(policy05)[20,7])`, semi-partial _r_ = `r printnum(r_policy05_pun, digits=2)`), or abortion (_b_ = `r printnum(parameterEstimates(policy06)[20,4], digits=2)`, 95% CI [`r printnum(parameterEstimates(policy06)[20,8], digits=2)`, `r printnum(parameterEstimates(policy06)[20,9], digits=2)`], _p_ = `r printp(parameterEstimates(policy06)[20,7])`, semi-partial _r_ = `r printnum(r_policy06_pun, digits=2)`). The punishment latent variable was also unrelated to a latent variable that captured progressive views on all these social policies combined ($\beta$ = `r printnum(standardizedSolution(policyLatent)[29,4])`, _b_ = `r printnum(parameterEstimates(policyLatent)[29,4])`, 95% CI [`r printnum(parameterEstimates(policyLatent)[29,8])`, `r printnum(parameterEstimates(policyLatent)[29,9])`], _p_ = `r printp(parameterEstimates(policyLatent)[29,7])`, semi-partial _r_ = `r printnum(sqrt(inspect(policyLatent, "r2")[["soc"]]))`).

@@

However, it is also possible that our norm-enforcing punishment and rule following tasks were poor measures of a general group conformist preference. As noted above, these tasks contain competitive and selfish elements that may explain their correlations with SDO rather than RWA. The tasks did not explicitly measure social norms [@Krupka2013] or allow players to interact over multiple rounds, limiting the potential for the emergence and enforcement of normative expectations. The tasks were also abstract and context-neutral: they did not specify who the players in the interaction were or contain any differentiation between in-groups and out-groups. This avoids conflating participants' general predisposition to conform with their identification with a particular group, but may fail to activate conformist preferences that depend on specific group contexts, such as socially conservative group identities [@Duckitt2009;@Everett2013]. Such an explanation would imply that authoritarians and social conservatives conform to social information in all contexts, but will only enforce and adhere to norms in specific group contexts. Future research could adapt the methods used here to further refine our understanding of the behavioural preferences underlying authoritarianism and social conservatism. In such work, it would be informative to relate preferences to specific ideological subdimensions, such as authoritarian aggression or authoritarian submission[@Duckitt2010b], in order to better map the behavioural constellation of the conformist dimension of political ideology.
